Interpretable Machine Learning

### 2019-01-09 chunks/story-wines.Rmd
continues analysis

@@

TODO: Show distribution of target

```{r load-data,include=FALSE}
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
source('../code/prepare-wine-data.R')
```

@@

TODO: Benchmark some models, report error

```{r learn}
library("mlr")
lrn.ranger = makeLearner("regr.ranger")
lrn.lm = makeLearner("regr.lm")
lrn.rpart = makeLearner("regr.rpart")

rdesc = makeResampleDesc("Holdout")

@@

bmr = benchmark(lrns, tsk, rdesc)

mod = mlr::train(lrn, tsk,  subset = seq(1, n, 2))
preds = predict(mod, task = tsk, subset = seq(2, n, 2))
performance(preds, measures = list(mae, mse))
pred  = Predictor$new(mod, data = wdat, y = "quality")

n_segs(pred)
ale_fanova(pred)

plot_all_ale = function(pred) {
  pls = lapply(setdiff(colnames(wdat), c("quality")), function(fname) {
    fe = FeatureEffect$new(pred, fname, method = "ale")
    res = fe$results
    colnames(res)[colnames(res) == fname] = ".feature"
    res$.feature.name = fname
    res$.feature = as.numeric(res$.feature)
    res$.feature.type = fe$feature.type
    res
  })

  res  = rbindlist(pls, use.names = TRUE)
  res$.feature = as.numeric(res$.feature)

  ggplot(res, aes(x = as.numeric(.feature), y= .ale)) +
    geom_line(aes(alpha = .feature.type)) +
    geom_point() +
    facet_wrap(".feature.name", scales = "free_x") +
    scale_alpha_discrete(guide = "none")
}

plot_all_ale(pred)

fi = FeatureImp$new(pred, loss = "mae", n.repetitions = 10)
fi$plot()

x.interest = wdat[20,]
sh = Shapley$new(pred, x.interest = x.interest)
plot(sh)

## Simpler model

lrn = makeLearner("regr.rpart", maxdepth = 2)
tsk = makeRegrTask(data = wdat, target = "quality")
mod = mlr::train(lrn, tsk,  subset = seq(1, n, 2))
preds = predict(mod, task = tsk, subset = seq(2, n, 2))
performance(preds, measures = list(mae, mse))
pred.tree  = Predictor$new(mod, data = wdat, y = "quality")

n_segs(pred.tree)
ale_fanova(pred.tree)

plot_all_ale(pred.tree)

mod$learner.model

fi = FeatureImp$new(pred, loss = "mae", n.repetitions = 10)
plot(fi)

plot_segs_feature(pred, feature.name = "alcohol", cat.mode = FALSE)
plot_segs_feature(pred, feature.name = "volatile.acidity", cat.mode = FALSE)
n_segs_feature(pred, "alcohol", cat.mode = TRUE)
n_segs_feature(pred, "sulphates", cat.mode = FALSE)
sh = Shapley$new(pred, x.interest = x.interest)
plot(sh)
li = LocalModel$new(pred, x.interest = x.interest)
plot(li)

lrn = makeLearner("regr.lm")
tsk = makeRegrTask(data = wdat, target = "quality")
mod = mlr::train(lrn, tsk,  subset = seq(1, n, 2))
preds = predict(mod, task = tsk, subset = seq(2, n, 2))
performance(preds, measures = list(mae, mse))
pred  = Predictor$new(mod, data = wdat, y = "quality")

library('mlr')
set.seed(42)
task = makeRegrTask(data = wine, target = 'quality')
lrn = makeLearner('regr.randomForest')
mod = train(lrn, task)
```

<div class="notes">
- Random Forest
- Target: Compensation
- All features that you found in the data
- Train, test, ship
</div>

@@

TODO: Benchmark some models, report error

Test: linear model, BRL, xgboost
Show table

TODO: Benchmark some models, report error

Test: linear model, BRL, xgboost
Show table

TODO: Show correlation between  predictions and true for test data

@@

TODO: Feature importances

@@

TODO: All effects

@@

TODO: Effect of alcohol

@@

## Shapley

TODO: Explain with Shapley

@@

>>>>

1 - fixed acidity
2 - volatile acidity
3 - citric acid
4 - residual sugar
5 - chlorides
6 - free sulfur dioxide
7 - total sulfur dioxide
8 - density
9 - pH
10 - sulphates
11 - alcohol
Output variable (based on sensory data):
12 - quality (score between 0 and 10)

Draw:
Volatile acidity as polish remover, pH as pH paper, alchohol as liqour bottle, chlorides as salt prinkler, citric acid as lemon, 
sulfur as mushroom?,

@@

```{r load-data,include=FALSE}
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
devtools::install_github("christophM/iml", ref = "feature-effects")
library('iml')
source('../code/prepare-wine-data.R')
```

```{r show-dist}
ggplot(wine) + 
  geom_bar(aes(x = quality)) + 
  scale_x_continuous("Wine Quality", labels = 1:10, breaks = 1:10)
```

@@

##

10x CV
Test linear regression model, decision tree, random forest

```{r benchmark}
library("mlr")
lrn.ranger = makeLearner("regr.ranger")
lrn.lm = makeLearner("regr.lm")
lrn.rpart = makeLearner("regr.rpart")

rdesc = cv5

@@

bmr = benchmark(lrns, tsk, rdesc, measures = list(mae))
bmr_tab = getBMRAggrPerformances(bmr, as.df = TRUE)
knitr::kable(bmr_tab[-1])
```

=> The random forest (ranger) is the best model.

```{r final-model}
mod = train(lrn.ranger, tsk)

set.seed(42)
sample_size = 500
wine_subset = wine[sample(1:nrow(wine), size = sample_size),]

pred = Predictor$new(mod, data = wine_subset, y = "quality")
```

@@

```{r check-model}

preds = pred$predict(wine)
preds$actual = wine$quality
ggplot(preds, aes(x = actual, y = .prediction, group = actual)) + 
  geom_violin(aes(fill = ..n..)) +
  scale_x_continuous("Actual quality (jittered)", 
    labels = 1:10, breaks = 1:10) + 
  scale_y_continuous("Predicted quality", labels = 1:10, breaks = 1:10) + 
  scale_fill_gradient(low = "white", high = "darkblue", guide = "none")
```

@@

```{r feature-importance}
imp = FeatureImp$new(pred, loss = "mae")
plot(imp)
```

@@

```{r feature-effects}
effs = FeatureEffects$new(pred)
plot(effs, ncols = 4)
```

@@

```{r feature-effect-alcohol}
eff = FeatureEffect$new(pred, "alcohol")
plot(eff)
```

## Rule of thumb what makes wine good?

```{r surrogate}
sbrl()
tree = TreeSurrogate$new(pred, maxdepth  = 2)
plot(tree)
```

@@

## Exceptionally bad wine

```{r bad-wine}
# find bad wine in data
predictions = pred$predict(wine)

min_pred = min(predictions)
worst = wine[which(predictions == min_pred),]

# explain with shapley
knitr::kable(t(worst))
# explain with cf?

min_pred

```

## Shapley Value

TODO: Draw features as players and within coalitions

## Shapley Value

```{r shapley}
shap = Shapley$new(pred, x.interest = worst)
plot(shap)
```

## Volatile Acidity

```{r volatile-acidity}
eff = FeatureEffect$new(pred, "volatile.acidity", method = "pdp+ice")
plot(eff)
```
## Volatile Acidity

‘Although the presence of high amounts of VA is considered undesirable, in some cases a touch of volatility is no bad thing,’ said Natasha Hughes MW in her guide to wine ‘flaws’.

```{r volatile-acidity}

# Get wines with low volatile acidity
predictions = pred$predict(wine_subset)

wine2 = wine_subset
wine2$volatile.acidity = wine2$volatile.acidity + 0.2
predictions2 = pred$predict(wine2)

plot(hist(predictions2$.prediction - predictions$.prediction))

wine_improve = which(predictions2$.prediction - predictions$.prediction > 0.3)

eff = FeatureEffect$new(pred, "volatile.acidity", method = "ice")

res = eff$results
res$improves = res$.id %in% wine_improve

ggplot(res) + 
  geom_line(aes(x = volatile.acidity, y = .y.hat, group = .id), alpha = 0.1) + 
  facet_wrap("improves")
```

@@

## Sulfur

At low levels reduction causes truffle, mushroom, radish, and green olive-like smells to wine. These have been accepted as positive traits and increase complexity in the wine. At higher levels, reduction is turns into a fault, giving wines a pungent smell of rotten egg, onion, garlic, cooked cabbage, canned corn, or even burnt rubber.

```{r sulfur}
```
